---
title: "Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
---
```{css my-style, echo = FALSE}

.chart-wrapper .chart-stage {
    overflow: auto;
}
```

```{r, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(plotly)
library(flexdashboard)

data("instacart")
```
Visualizations {data-icon="fa-signal"}
===================================== 
### Chart A

```{r, warning = FALSE, message = FALSE, echo = FALSE}
frozen_instacart = instacart %>% 
  filter(department == 'frozen') %>% 
  add_count(product_name) %>% 
  filter(n >= 400)

frozen_instacart %>% 
  count(product_name) %>%
  mutate(product_name = as.character(product_name)) %>%
  mutate(product_name = fct_reorder(product_name, n)) %>%
    plot_ly(
      x = ~product_name, y = ~n, color = ~product_name, type = "bar", colors = "viridis") %>%
  layout(title = 'What Frozen Foods New Yorkers are Ordering?', xaxis = list(title = "Product"), yaxis = list(title = "Count")) %>% hide_legend()
```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r, echo = FALSE}
frozen_instacart %>% 
  mutate(product_name = fct_reorder(product_name, days_since_prior_order)) %>% 
  plot_ly(y = ~days_since_prior_order, color = ~product_name, type = "box", colors = "viridis") %>%
  layout(
    title = 'How Often Are New Yorkers Ordering Their Favorite Frozen Foods?', xaxis = list(title = "Product"), yaxis = list(title = "Days Between Orders")) %>% hide_legend()
```

### Chart C

```{r, warning = FALSE, message = FALSE, echo = FALSE}
frozen_instacart_time = frozen_instacart %>% group_by(product_name, order_hour_of_day) %>% summarize(count = n())

plot_ly(data = frozen_instacart_time, x = ~order_hour_of_day, y = ~count, color = ~product_name, mode = 'lines') %>%
  layout(
    title = 'When Are New Yorkers Ordering Their Favorite Frozen Foods?', xaxis = list(title = "Hour (Military Time)"), yaxis = list(title = "Number of Products Ordered"))

```

About {data-icon="fa-info"}
=====================================   
### Introduction

These dashboards are analyzing “The Instacart Online Grocery Shopping Dataset 2017”, Accessed from https://www.instacart.com/datasets/grocery-shopping-2017 on November 1, 2022.

The aim is to identify frequently ordered frozen products and discover any purchase patterns.

Note: this analysis is only looking at products that had over 400 purchases in the timeframe.

